---
title: "Getting Started with zoomstudentengagement"
author: "zoomstudentengagement package"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with zoomstudentengagement}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

```{r setup}
library(zoomstudentengagement)
library(dplyr)
library(ggplot2)
```

# Getting Started with zoomstudentengagement

The `zoomstudentengagement` package helps instructors analyze student engagement from Zoom transcripts, with a particular focus on participation equity. This vignette will get you started with the basic workflow.

## Installation

Install the package from GitHub:

```{r eval = FALSE}
devtools::install_github("revgizmo/zoomstudentengagement")
```

## Quick Start

Here's a minimal example to get you started:

```{r quick-start}
# Load a sample transcript
transcript_file <- system.file(
  "extdata/transcripts/GMT20240124-202901_Recording.transcript.vtt",
  package = "zoomstudentengagement"
)

# Process the transcript
processed_transcript <- process_zoom_transcript(
  transcript_file_path = transcript_file,
  consolidate_comments = TRUE,
  add_dead_air = TRUE
)

# Calculate summary metrics
summary_metrics <- summarize_transcript_metrics(
  transcript_file_path = transcript_file,
  names_exclude = c("dead_air")
)

# View the results
head(summary_metrics)
```

## What the Package Does

The `zoomstudentengagement` package provides tools for:

1. **Loading and Processing Zoom Transcripts**: Convert Zoom VTT files into analyzable data
2. **Calculating Engagement Metrics**: Measure participation by speaker/student
3. **Name Matching and Cleaning**: Match transcript names to student rosters
4. **Visualization**: Create plots to analyze participation patterns
5. **Reporting**: Generate individual student reports

## Basic Workflow

The typical workflow involves:

1. **Setup**: Configure your analysis parameters
2. **Load Transcripts**: Import and process Zoom transcript files
3. **Load Roster**: Import student enrollment data
4. **Clean Names**: Match transcript names to student records
5. **Analyze**: Calculate metrics and create visualizations
6. **Report**: Generate insights and reports

## Next Steps

- **For transcript processing**: See the [Transcript Processing](transcript-processing.html) vignette
- **For roster management**: See the [Roster Cleaning](roster-cleaning.html) vignette  
- **For visualization**: See the [Plotting and Analysis](plotting.html) vignette
- **For advanced workflows**: See the [Session Mapping](session-mapping.html) vignette

## Getting Help

- Check the [package documentation](https://revgizmo.github.io/zoomstudentengagement/)
- Review the [PROJECT.md](https://github.com/revgizmo/zoomstudentengagement_cursor/blob/main/PROJECT.md) for current development status
- Report issues on [GitHub](https://github.com/revgizmo/zoomstudentengagement_cursor/issues) 