---
title: "Plotting and Analysis with zoomstudentengagement"
author: "zoomstudentengagement package"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plotting and Analysis with zoomstudentengagement}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6
)
```

```{r setup}
library(zoomstudentengagement)
library(dplyr)
library(ggplot2)
```

# Plotting and Analysis

This vignette shows how to create visualizations and analyze student engagement patterns using the `zoomstudentengagement` package.

## Preparing Data for Analysis

### Loading Sample Data

First, let's prepare some sample data for analysis:

```{r prepare-data}
# Load sample data
data_folder <- system.file("extdata", package = "zoomstudentengagement")

# Load roster
roster_df <- load_roster(
  data_folder = data_folder,
  roster_file = "roster.csv"
)

# Load transcript metrics
transcript_file <- system.file(
  "extdata/transcripts/GMT20240124-202901_Recording.transcript.vtt",
  package = "zoomstudentengagement"
)

transcripts_metrics_df <- summarize_transcript_metrics(
  transcript_file_path = transcript_file,
  names_exclude = c("dead_air")
)

# Create session data (simplified for example)
transcripts_list_df <- data.frame(
  transcript_file = "GMT20240124-202901_Recording.transcript.vtt",
  session_date = as.Date("2024-01-24"),
  section = "LTF.201.1"
)

roster_small_df <- make_roster_small(roster_df)
roster_sessions <- make_student_roster_sessions(
  transcripts_list_df,
  roster_small_df
)

# Create clean names
clean_names_df <- make_clean_names_df(
  data_folder = data_folder,
  section_names_lookup_file = "section_names_lookup.csv",
  transcripts_metrics_df,
  roster_sessions
)

# Create summary dataframes
transcripts_session_summary_df <- make_transcripts_session_summary_df(clean_names_df)
transcripts_summary_df <- make_transcripts_summary_df(transcripts_session_summary_df)
```

## Basic Visualization

### Plotting by Metric

The `plot_users_by_metric()` function creates visualizations for different engagement metrics:

```{r basic-plots}
# Plot session count
plot_users_by_metric(transcripts_summary_df, metric = "session_ct")
```

```{r comment-count}
# Plot comment count
plot_users_by_metric(transcripts_summary_df, metric = "n")
```

```{r duration}
# Plot speaking duration
plot_users_by_metric(transcripts_summary_df, metric = "duration")
```

```{r wordcount}
# Plot word count
plot_users_by_metric(transcripts_summary_df, metric = "wordcount")
```

## Understanding the Metrics

### Available Metrics

The package provides several engagement metrics:

- **session_ct**: Number of sessions attended
- **n**: Number of comments made
- **perc_n**: Percentage of total comments
- **duration**: Total speaking time (seconds)
- **perc_duration**: Percentage of total speaking time
- **wordcount**: Total words spoken
- **perc_wordcount**: Percentage of total words
- **wpm**: Words per minute

### Percentage Metrics

Visualize relative participation:

```{r percentage-metrics}
# Plot percentage of comments
plot_users_by_metric(transcripts_summary_df, metric = "perc_n")

# Plot percentage of speaking time
plot_users_by_metric(transcripts_summary_df, metric = "perc_duration")

# Plot percentage of words
plot_users_by_metric(transcripts_summary_df, metric = "perc_wordcount")
```

## Students-Only Analysis

### Filtering for Enrolled Students

Focus analysis on enrolled students only:

```{r students-only}
# Create students-only summary
students_only_summary <- make_students_only_transcripts_summary_df(
  transcripts_session_summary_df
)

# Plot students-only metrics
plot_users_by_metric(students_only_summary, metric = "session_ct")
```

### Masked Names for Privacy

Use masked names for privacy-conscious analysis:

```{r masked-plots}
# Plot with masked names
plot_users_masked_section_by_metric(
  df = students_only_summary,
  metric = "n"
)

plot_users_masked_section_by_metric(
  df = students_only_summary,
  metric = "duration"
)
```

## Advanced Analysis

### Custom Filtering

Filter data for specific analysis:

```{r custom-filtering}
# Filter for specific sections
section_data <- transcripts_summary_df %>%
  filter(section == "LTF.201.1")

# Plot filtered data
plot_users_by_metric(section_data, metric = "wpm")
```

### Comparative Analysis

Compare different metrics:

```{r comparative}
# Create comparison plots
par(mfrow = c(2, 2))
plot_users_by_metric(transcripts_summary_df, metric = "n")
plot_users_by_metric(transcripts_summary_df, metric = "duration")
plot_users_by_metric(transcripts_summary_df, metric = "wordcount")
plot_users_by_metric(transcripts_summary_df, metric = "wpm")
```

## Interpreting the Results

### Participation Equity

Look for patterns in participation:

```{r equity-analysis}
# Analyze participation distribution
participation_summary <- transcripts_summary_df %>%
  group_by(section) %>%
  summarise(
    total_students = n(),
    active_students = sum(n > 0),
    avg_comments = mean(n),
    median_comments = median(n),
    participation_rate = active_students / total_students
  )

participation_summary
```

### Engagement Patterns

Identify different types of engagement:

```{r patterns}
# Categorize students by engagement type
engagement_categories <- transcripts_summary_df %>%
  mutate(
    engagement_type = case_when(
      n == 0 ~ "No participation",
      n <= 2 ~ "Low participation", 
      n <= 5 ~ "Moderate participation",
      TRUE ~ "High participation"
    )
  ) %>%
  count(engagement_type)

engagement_categories
```

## Custom Visualizations

### Creating Your Own Plots

Use the data to create custom visualizations:

```{r custom-plots}
# Custom participation distribution
ggplot(transcripts_summary_df, aes(x = n)) +
  geom_histogram(binwidth = 1, fill = "steelblue", alpha = 0.7) +
  labs(
    title = "Distribution of Comment Counts",
    x = "Number of Comments",
    y = "Number of Students"
  ) +
  theme_minimal()

# Custom duration vs word count
ggplot(transcripts_summary_df, aes(x = duration, y = wordcount)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Speaking Duration vs Word Count",
    x = "Duration (seconds)",
    y = "Word Count"
  ) +
  theme_minimal()
```

## Best Practices

### Choosing the Right Metric

- **For attendance**: Use `session_ct`
- **For participation frequency**: Use `n` or `perc_n`
- **For speaking time**: Use `duration` or `perc_duration`
- **For content contribution**: Use `wordcount` or `perc_wordcount`
- **For speaking pace**: Use `wpm`

### Privacy Considerations

- Use masked names when sharing results
- Consider aggregating data for small groups
- Be mindful of FERPA compliance requirements

### Visualization Tips

- Use consistent color schemes
- Include clear titles and labels
- Consider log scales for skewed distributions
- Add context (e.g., class averages, benchmarks)

## Next Steps

- **For session mapping**: See the [Session Mapping](session-mapping.html) vignette
- **For student reports**: See the [Student Reports](student-reports.html) vignette
- **For troubleshooting**: See the [Troubleshooting Guide](troubleshooting.html) vignette 