---
title: "Plotting and Analysis with zoomstudentengagement"
author: "zoomstudentengagement package"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plotting and Analysis with zoomstudentengagement}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6
)
```

```{r setup}
library(zoomstudentengagement)
library(dplyr)
library(ggplot2)
```

# Plotting and Analysis

This vignette shows how to create visualizations and analyze student engagement patterns using the `zoomstudentengagement` package.

## Preparing Data for Analysis

### Loading Sample Data

First, let's prepare some sample data for analysis:

```{r prepare-data}
# Create sample data for demonstration
# In practice, you would load actual transcript and roster data

# Sample transcript metrics
transcripts_metrics_df <- tibble::tibble(
  name = c("Alice Johnson", "Bob Smith", "Carol Davis", "David Wilson", "Eva Brown"),
  n = c(8, 12, 5, 15, 3),
  duration = c(45.2, 67.8, 23.1, 89.4, 12.5),
  wordcount = c(1200, 1800, 650, 2200, 400),
  comments = list("Good point", "Interesting question", "I agree", "Follow-up question", "Brief comment"),
  n_perc = c(18.6, 27.9, 11.6, 34.9, 7.0),
  duration_perc = c(19.2, 28.8, 9.8, 38.0, 5.3),
  wordcount_perc = c(19.2, 28.8, 10.4, 35.2, 6.4),
  wpm = c(26.5, 26.6, 28.1, 24.6, 32.0),
  course_section = "LTF.201.1",
  course = 201,
  section = 1,
  day = "Thursday",
  time = "18:30",
  name_raw = name,
  start_time_local = as.POSIXct("2024-01-24 18:30:00", tz = "America/Los_Angeles"),
  dept = "LTF",
  session_num = 1
)

# Sample roster sessions
roster_sessions <- tibble::tibble(
  student_id = c("12345", "12346", "12347", "12348", "12349"),
  first_last = c("Alice Johnson", "Bob Smith", "Carol Davis", "David Wilson", "Eva Brown"),
  preferred_name = c("Alice", "Bob", "Carol", "David", "Eva"),
  dept = "LTF",
  course = 201,
  section = 1,
  session_num = 1,
  start_time_local = as.POSIXct("2024-01-24 18:30:00", tz = "America/Los_Angeles"),
  course_section = "LTF.201.1"
)

# Create clean names dataframe
# Use a temporary directory since we're working with sample data
temp_dir <- tempdir()
clean_names_df <- make_clean_names_df(
  data_folder = temp_dir,
  section_names_lookup_file = "section_names_lookup.csv",
  transcripts_metrics_df,
  roster_sessions
)

# Create summary dataframes
transcripts_session_summary_df <- make_transcripts_session_summary_df(clean_names_df)
transcripts_summary_df <- make_transcripts_summary_df(transcripts_session_summary_df)

# View the summary data
head(transcripts_summary_df)
```

## Basic Visualization

### Plotting by Metric

Use the unified `plot_users()` function to create visualizations for different engagement metrics:

```{r basic-plots}
# Plot session count
plot_users(transcripts_summary_df, metric = "session_ct", facet_by = "section", mask_by = "name")
```

```{r comment-count}
# Plot comment count
plot_users(transcripts_summary_df, metric = "n", facet_by = "section", mask_by = "name")
```

```{r duration}
# Plot speaking duration
plot_users(transcripts_summary_df, metric = "duration", facet_by = "section", mask_by = "name")
```

```{r wordcount}
# Plot word count
plot_users(transcripts_summary_df, metric = "wordcount", facet_by = "section", mask_by = "name")
```

## Understanding the Metrics

### Available Metrics

The package provides several engagement metrics:

- **session_ct**: Number of sessions attended
- **n**: Number of comments made
- **n_perc**: Percentage of total comments
- **duration**: Total speaking time (seconds)
- **duration_perc**: Percentage of total speaking time
- **wordcount**: Total words spoken
- **wordcount_perc**: Percentage of total words
- **wpm**: Words per minute

### Percentage Metrics

Visualize relative participation:

```{r percentage-metrics}
# Plot percentage of comments
plot_users(transcripts_summary_df, metric = "n_perc", facet_by = "section", mask_by = "name")

# Plot percentage of speaking time
plot_users(transcripts_summary_df, metric = "duration_perc", facet_by = "section", mask_by = "name")

# Plot percentage of words
plot_users(transcripts_summary_df, metric = "wordcount_perc", facet_by = "section", mask_by = "name")
```

## Students-Only Analysis

### Filtering for Enrolled Students

Focus analysis on enrolled students only:

```{r students-only}
# Create students-only summary
students_only_summary <- make_students_only_transcripts_summary_df(
  transcripts_session_summary_df
)

# Plot students-only metrics
plot_users(students_only_summary, metric = "session_ct", facet_by = "section", mask_by = "name")
```

### Masked Names for Privacy

Use rank-based masked labels for privacy-conscious analysis:

```{r masked-plots}
# Plot with rank-masked names
plot_users(
  data = students_only_summary,
  metric = "n",
  facet_by = "section",
  mask_by = "rank"
)

plot_users(
  data = students_only_summary,
  metric = "duration",
  facet_by = "section",
  mask_by = "rank"
)
```

## Advanced Analysis

### Custom Filtering

Filter data for specific analysis:

```{r custom-filtering}
# Filter for specific sections
section_data <- transcripts_summary_df %>%
  filter(section == 1) # Use the actual section number from our sample data

# Plot filtered data
plot_users(section_data, metric = "wpm", facet_by = "section", mask_by = "name")
```

### Comparative Analysis

Compare different metrics:

```{r comparative}
# Create comparison plots
par(mfrow = c(2, 2))
plot_users(transcripts_summary_df, metric = "n", facet_by = "section", mask_by = "name")
plot_users(transcripts_summary_df, metric = "duration", facet_by = "section", mask_by = "name")
plot_users(transcripts_summary_df, metric = "wordcount", facet_by = "section", mask_by = "name")
plot_users(transcripts_summary_df, metric = "wpm", facet_by = "section", mask_by = "name")
```

## Interpreting the Results

### Participation Equity

Look for patterns in participation:

```{r equity-analysis}
# Analyze participation distribution
participation_summary <- transcripts_summary_df %>%
  group_by(section) %>%
  summarise(
    total_students = n(),
    active_students = sum(n > 0),
    avg_comments = mean(n),
    median_comments = median(n),
    participation_rate = active_students / total_students
  )

participation_summary
```

## Custom Visualizations

### Creating Your Own Plots

Use the data to create custom visualizations:

```{r custom-plots}
# Custom participation distribution
ggplot(transcripts_summary_df, aes(x = n)) +
  geom_histogram(binwidth = 1, fill = "steelblue", alpha = 0.7) +
  labs(
    title = "Distribution of Comment Counts",
    x = "Number of Comments",
    y = "Number of Students"
  ) +
  theme_minimal()

# Custom duration vs word count
ggplot(transcripts_summary_df, aes(x = duration, y = wordcount)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Speaking Duration vs Word Count",
    x = "Duration (seconds)",
    y = "Word Count"
  ) +
  theme_minimal()
```

## Best Practices

### Choosing the Right Metric

- **For attendance**: Use `session_ct`
- **For participation frequency**: Use `n` or `n_perc`
- **For speaking time**: Use `duration` or `duration_perc`
- **For content contribution**: Use `wordcount` or `wordcount_perc`
- **For speaking pace**: Use `wpm`

### Privacy Considerations

- Use masked names when sharing results
- Consider aggregating data for small groups
- Be mindful of FERPA compliance requirements

### Visualization Tips

- Use consistent color schemes
- Include clear titles and labels
- Consider log scales for skewed distributions
- Add context (e.g., class averages, benchmarks)

## Next Steps

- **For session mapping**: See the [Session Mapping](session-mapping.html) vignette
- **For student reports**: See the [Student Reports](student-reports.html) vignette
- **For troubleshooting**: See the [Troubleshooting Guide](troubleshooting.html) vignette 
