---
title: "Roster Management and Name Cleaning"
author: "zoomstudentengagement package"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Roster Management and Name Cleaning}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

```{r setup}
library(zoomstudentengagement)
library(dplyr)
library(ggplot2)
```

# Roster Management and Name Cleaning

This vignette shows how to manage student rosters and clean name mismatches between Zoom transcripts and official enrollment data.

## Loading Student Rosters

### Basic Roster Loading

Load a student roster from a CSV file:

```{r load-roster}
# Load the sample roster
data_folder <- system.file("extdata", package = "zoomstudentengagement")

roster_df <- load_roster(
  data_folder = data_folder,
  roster_file = "roster.csv"
)

# View the roster structure
head(roster_df)
```

### Creating Section Summaries

Generate summaries by course section:

```{r sections}
# Create sections summary
sections_df <- make_sections_df(roster_df)

# View sections
sections_df
```

### Creating Student Lists

Create a simplified student list for analysis:

```{r roster-small}
# Create simplified roster
roster_small_df <- make_roster_small(roster_df)

# View simplified roster
head(roster_small_df)
```

## Working with Transcript Data

### Loading Transcript Metrics

First, let's load some transcript data to work with:

```{r load-transcripts}
# Create sample transcript metrics for demonstration
# In practice, you would use summarize_transcript_metrics() on actual transcript files
transcripts_metrics_df <- tibble::tibble(
  name = c("Conor Healy", "Dr. Melissa Ko", "Ryan Sloan"),
  n = c(15, 8, 12),
  duration = c(45.2, 23.1, 38.7),
  wordcount = c(1200, 650, 950),
  comments = list("Good discussion", "Excellent points", "Interesting question"),
  n_perc = c(25.0, 13.3, 20.0),
  duration_perc = c(25.0, 13.3, 20.0),
  wordcount_perc = c(25.0, 13.3, 20.0),
  wpm = c(26.5, 28.1, 24.6),
  course_section = "23.24",
  course = 23,
  section = 24,
  day = "Thursday",
  time = "18:30",
  name_raw = name,
  start_time_local = as.POSIXct("2024-01-24 18:30:00", tz = "America/Los_Angeles"),
  dept = "LTF",
  session_num = 1
)

# View transcript metrics
head(transcripts_metrics_df)
```

### Creating Session Rosters

Create a roster with entries for each recorded session:

```{r session-roster}
# Create a simplified session mapping for demonstration
# In practice, you would use the full data loading functions
transcripts_list_df <- tibble::tibble(
  dept = "LTF",
  course = 23,
  section = 24,
  session_num = 1,
  start_time_local = as.POSIXct("2024-01-24 18:30:00", tz = "America/Los_Angeles"),
  course_section = "23.24"
)

# Create roster sessions
roster_sessions <- make_student_roster_sessions(
  transcripts_list_df,
  roster_small_df
)

# View roster sessions
head(roster_sessions)
```

## Name Cleaning Process

### Initial Name Matching

Join transcript names with roster data:

```{r initial-matching}
# Create clean names dataframe
clean_names_df <- make_clean_names_df(
  data_folder = data_folder,
  section_names_lookup_file = "section_names_lookup.csv",
  transcripts_metrics_df,
  roster_sessions
)

# View initial matching results
head(clean_names_df)
```

### Identifying Names to Clean

Find names that need manual cleaning:

```{r names-to-clean}
# Find names that need cleaning
names_to_clean <- make_names_to_clean_df(clean_names_df)

# View names needing attention
names_to_clean
```

### Writing Names Lookup

Create a lookup file for manual editing:

```{r write-lookup}
# Write section names lookup
write_section_names_lookup(
  clean_names_df,
  data_folder = data_folder,
  section_names_lookup_file = "section_names_lookup.csv"
)
```

## Manual Name Cleaning

### Understanding the Process

The name cleaning process involves:

1. **Automatic Matching**: The package attempts to match transcript names to roster names
2. **Manual Review**: Names that don't match automatically are flagged for review
3. **Lookup File**: A CSV file is created for manual editing
4. **Iterative Process**: Repeat until all names are properly matched

### Common Name Issues

Typical issues you might encounter:

- **Nicknames**: "Mike" vs "Michael"
- **Display Names**: "Prof. Smith" vs "Dr. Smith"  
- **Guest Speakers**: Names not in the roster
- **Technical Issues**: "Unknown" or "User123"

### Editing the Lookup File

The lookup file contains:

- **transcript_name**: Name as it appears in the transcript
- **preferred_name**: How you want it to appear in analysis
- **formal_name**: Official name from roster
- **student_id**: Student identifier
- **section**: Course section

## Advanced Name Cleaning

### Using Preferred Names

The package supports preferred names from the roster:

```{r preferred-names}
# Check if roster has preferred names
if ("preferred_name" %in% colnames(roster_df)) {
  cat("Roster includes preferred names\n")
  head(roster_df[, c("first_last", "preferred_name")])
} else {
  cat("Roster uses formal names only\n")
}
```

### Handling Special Cases

For special cases like guest speakers:

```{r special-cases}
# Example: Mark guest speakers
# In the lookup file, set:
# transcript_name: "Guest Speaker"
# preferred_name: "Guest"
# formal_name: "Guest Speaker"
# student_id: "GUEST001"
# section: "LTF.201.1"
```

## Iterative Cleaning Process

### Step-by-Step Workflow

1. **Run initial matching**:
   ```r
   clean_names_df <- make_clean_names_df(...)
   ```

2. **Check for unmatched names**:
   ```r
   names_to_clean <- make_names_to_clean_df(clean_names_df)
   ```

3. **Edit the lookup file** manually

4. **Re-run matching** with updated lookup

5. **Repeat** until all names are matched

### Validation

Check your cleaning results:

```{r validation}
# After cleaning, check results
final_clean_names <- make_clean_names_df(
  data_folder = data_folder,
  section_names_lookup_file = "section_names_lookup.csv",
  transcripts_metrics_df,
  roster_sessions
)

# Check for any remaining unmatched names
remaining_unmatched <- make_names_to_clean_df(final_clean_names)
if (nrow(remaining_unmatched) == 0) {
  cat("All names successfully matched!\n")
} else {
  cat("Still have", nrow(remaining_unmatched), "names to clean\n")
}
```

## Next Steps

- **For analysis**: See the [Plotting and Analysis](plotting.html) vignette
- **For session mapping**: See the [Session Mapping](session-mapping.html) vignette
- **For student reports**: See the [Student Reports](student-reports.html) vignette 
