---
title: "Whole Game: Complete Workflow for New Instructors"
author: "zoomstudentengagement package"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Whole Game: Complete Workflow for New Instructors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6
)
```

```{r load-packages}
library(zoomstudentengagement)
library(dplyr)
library(ggplot2)
library(readr)
library(tibble)
```

# Whole Game: Complete Workflow for New Instructors

This vignette demonstrates the complete workflow for analyzing student engagement in Zoom sessions. It's designed for new instructors who want to understand participation patterns and promote equitable engagement in their classes.

## Scenario

You're a new instructor who has:
- Zoom transcripts for several sessions of your CS 101 course
- Multiple Zoom CSV files saved as you added transcripts to your transcript folder
- A roster of enrolled students that needs to be aligned with Zoom transcript names
- A goal to understand participation patterns and promote equitable engagement

## Step 1: Understanding Your Data

First, let's explore what data you have available:

```{r explore-data}
# Check what transcript files are available
transcript_files <- list.files(
  system.file("extdata/transcripts", package = "zoomstudentengagement"),
  pattern = "\\.vtt$",
  full.names = TRUE
)

cat("Available transcript files:\n")
basename(transcript_files)

# Check what other data files are available
cat("\nAvailable data files:\n")
list.files(system.file("extdata", package = "zoomstudentengagement"))
```

## Step 2: Loading Your Student Roster

Start by loading your student roster:

```{r load-roster}
# Load the student roster
roster <- load_roster(
  data_folder = system.file("extdata", package = "zoomstudentengagement"),
  roster_file = "roster.csv"
)

# View the roster structure
head(roster)
cat("\nRoster dimensions:", nrow(roster), "students\n")
```

## Step 3: Loading and Processing Transcript Files

Now let's load and process your Zoom transcripts:

```{r load-transcripts}
# Load a single transcript to see the structure
transcript_file <- system.file(
  "extdata/transcripts/GMT20240124-202901_Recording.transcript.vtt",
  package = "zoomstudentengagement"
)

# Load and process the transcript
transcript_data <- load_zoom_transcript(transcript_file)

# View the structure
head(transcript_data)
cat("\nTranscript dimensions:", nrow(transcript_data), "utterances\n")
```

## Step 4: Processing Multiple Sessions

If you have multiple sessions, you can process them all at once:

```{r process-multiple}
# Get all transcript files
transcript_files <- list.files(
  system.file("extdata/transcripts", package = "zoomstudentengagement"),
  pattern = "\\.vtt$",
  full.names = TRUE
)

# Process all transcripts
all_transcripts <- lapply(transcript_files, function(file) {
  cat("Processing:", basename(file), "\n")
  load_zoom_transcript(file)
})

# Combine all transcripts
combined_transcripts <- dplyr::bind_rows(all_transcripts, .id = "session")

cat("\nCombined data dimensions:", nrow(combined_transcripts), "utterances\n")
```

## Step 5: Calculating Engagement Metrics

Now let's calculate engagement metrics for each student:

```{r calculate-metrics}
# Calculate metrics for all transcript files
metrics <- summarize_transcript_files(
  transcript_file_names = basename(transcript_files),
  data_folder = system.file("extdata", package = "zoomstudentengagement"),
  transcripts_folder = "transcripts",
  names_to_exclude = c("dead_air", "Unknown")
)

# View the metrics
head(metrics)
```

## Step 6: Matching Student Names

This is often the most challenging part - matching Zoom names to your roster:

```{r name-matching}
# First, let's see what names appear in the transcripts
transcript_names <- unique(metrics$name)
cat("Names in transcripts:\n")
print(transcript_names)

# Compare with roster names
roster_names <- unique(roster$preferred_name)
cat("\nNames in roster:\n")
print(roster_names)

# For this example, let's create a simple manual mapping
# In practice, you would use the more advanced name matching functions
name_mapping <- data.frame(
  name_to_clean = transcript_names,
  clean_name = ifelse(transcript_names %in% roster_names, transcript_names, NA_character_),
  stringsAsFactors = FALSE
)

# View the mapping
print(name_mapping)
```

## Step 7: Handling Preferred Names

Students often use preferred names in Zoom that don't match their official roster:

```{r preferred-names}
# Identify names that need manual mapping
unmatched_names <- name_mapping %>%
  filter(is.na(clean_name)) %>%
  pull(name_to_clean)

cat("Names needing manual mapping:\n")
print(unmatched_names)

# For this example, let's create some manual mappings
manual_mappings <- data.frame(
  name_to_clean = c("Bob", "Sally", "Mike"),
  clean_name = c("Robert Smith", "Sarah Johnson", "Michael Brown"),
  notes = c("Uses nickname", "Preferred name", "Uses nickname"),
  stringsAsFactors = FALSE
)

# Update the name mapping
name_mapping <- bind_rows(
  name_mapping %>% filter(!is.na(clean_name)),
  manual_mappings
)
```

## Step 8: Creating Clean Engagement Data

Now let's apply the name mappings to get clean engagement data:

```{r clean-engagement}
# Apply name mappings to metrics
clean_metrics <- metrics %>%
  left_join(name_mapping, by = c("name" = "name_to_clean")) %>%
  mutate(
    student_name = coalesce(clean_name, name),
    is_matched = !is.na(clean_name)
  )

# View the results
head(clean_metrics)

# Summary of matching success
matching_summary <- clean_metrics %>%
  group_by(is_matched) %>%
  summarise(
    count = n(),
    percentage = n() / nrow(clean_metrics) * 100
  )

print(matching_summary)
```

## Step 9: Analyzing Participation Patterns

Now let's analyze participation patterns to understand engagement:

```{r analyze-patterns}
# Create a summary by student
student_summary <- clean_metrics %>%
  group_by(student_name) %>%
  summarise(
    total_utterances = sum(n, na.rm = TRUE),
    total_duration = sum(duration, na.rm = TRUE),
    total_words = sum(wordcount, na.rm = TRUE),
    avg_words_per_minute = mean(wpm, na.rm = TRUE),
    participation_rate = total_utterances / nrow(clean_metrics) * 100
  ) %>%
  arrange(desc(total_utterances))

# View the summary
head(student_summary, 10)
```

## Step 10: Visualizing Participation

Create visualizations to understand participation patterns:

```{r visualize-participation}
# Plot participation by utterance count
ggplot(student_summary, aes(x = reorder(student_name, total_utterances), y = total_utterances)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  labs(
    title = "Student Participation by Utterance Count",
    x = "Student Name",
    y = "Number of Utterances"
  ) +
  theme_minimal()

# Plot participation by duration
ggplot(student_summary, aes(x = reorder(student_name, total_duration), y = total_duration / 60)) +
  geom_bar(stat = "identity", fill = "darkgreen") +
  coord_flip() +
  labs(
    title = "Student Participation by Speaking Time",
    x = "Student Name",
    y = "Speaking Time (minutes)"
  ) +
  theme_minimal()
```

## Step 11: Identifying Participation Gaps

Look for students who might need encouragement to participate:

```{r identify-gaps}
# Identify students with low participation
low_participation <- student_summary %>%
  filter(total_utterances < median(total_utterances, na.rm = TRUE)) %>%
  arrange(total_utterances)

cat("Students with below-median participation:\n")
print(low_participation[, c("student_name", "total_utterances", "total_duration")])

# Calculate participation equity metrics
participation_stats <- student_summary %>%
  summarise(
    mean_utterances = mean(total_utterances, na.rm = TRUE),
    median_utterances = median(total_utterances, na.rm = TRUE),
    sd_utterances = sd(total_utterances, na.rm = TRUE),
    gini_coefficient = (2 * sum(rank(total_utterances) * total_utterances) /
      (n() * sum(total_utterances))) - (n() + 1) / n()
  )

cat("\nParticipation Statistics:\n")
print(participation_stats)
```

## Step 12: Creating Actionable Insights

Generate insights that can help improve equitable participation:

```{r actionable-insights}
# Create participation categories
participation_categories <- student_summary %>%
  mutate(
    participation_level = case_when(
      total_utterances >= quantile(total_utterances, 0.75, na.rm = TRUE) ~ "High",
      total_utterances >= quantile(total_utterances, 0.25, na.rm = TRUE) ~ "Medium",
      TRUE ~ "Low"
    )
  )

# Summary by participation level
level_summary <- participation_categories %>%
  group_by(participation_level) %>%
  summarise(
    count = n(),
    percentage = n() / nrow(participation_categories) * 100
  )

cat("Participation Level Distribution:\n")
print(level_summary)

# Recommendations based on analysis
cat("\n=== RECOMMENDATIONS FOR EQUITABLE PARTICIPATION ===\n")
cat("1. Students with low participation may benefit from:\n")
cat("   - Direct invitations to contribute\n")
cat("   - Smaller group discussions\n")
cat("   - Alternative participation methods (chat, polls)\n\n")

cat("2. Consider implementing:\n")
cat("   - Structured discussion protocols\n")
cat("   - Think-pair-share activities\n")
cat("   - Anonymous participation options\n\n")

cat("3. Monitor participation patterns over time to:\n")
cat("   - Track improvement in engagement\n")
cat("   - Identify effective interventions\n")
cat("   - Ensure all students feel included\n")
```

## Step 13: Saving Your Analysis

Save your processed data for future reference:

```{r save-analysis}
# Save the clean metrics
write_csv(clean_metrics, "clean_engagement_metrics.csv")

# Save the student summary
write_csv(student_summary, "student_participation_summary.csv")

# Save the name mappings for future use
write_csv(name_mapping, "name_mappings.csv")

cat("Analysis files saved:\n")
cat("- clean_engagement_metrics.csv\n")
cat("- student_participation_summary.csv\n")
cat("- name_mappings.csv\n")
```

## Best Practices and Tips

### Data Organization
- Keep your transcript files organized by date/session
- Maintain a consistent naming convention
- Back up your original data files
- Document any manual name mappings

### Ethical Considerations
- Use this data to promote equitable participation, not surveillance
- Focus on group patterns rather than individual performance
- Respect student privacy and preferences
- Share insights constructively with students when appropriate

### Common Pitfalls to Avoid
- Don't assume low participation means disengagement
- Consider cultural and personal communication preferences
- Remember that quality of participation matters more than quantity
- Avoid making assumptions about student abilities based on participation

### Interpreting Results
- Look for patterns, not individual judgments
- Consider context (class size, format, topic)
- Use data to inform teaching strategies, not student evaluation
- Focus on creating inclusive learning environments

## Next Steps

This analysis provides a foundation for understanding participation patterns. Consider:

1. **Regular monitoring**: Track participation patterns over the semester
2. **Intervention strategies**: Implement targeted approaches for low-participation students
3. **Student feedback**: Ask students about their participation preferences
4. **Pedagogical adjustments**: Modify teaching strategies based on insights
5. **Continuous improvement**: Refine your analysis approach over time

Remember: The goal is to create an inclusive learning environment where all students feel comfortable and encouraged to participate in ways that work for them. 
