---
title: "Whole Game: Real-World Workflow for Instructors"
author: "zoomstudentengagement package"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6,
  warning = FALSE,
  message = FALSE
)
```

```{r load-packages}
library(zoomstudentengagement)
library(dplyr)
library(ggplot2)
library(readr)
library(tibble)
library(lubridate)
library(jsonlite)
```

# Whole Game: Real-World Workflow for Instructors

This document demonstrates the complete workflow for analyzing student engagement in Zoom sessions using your actual data. It's designed to work with the real-world testing environment.

## Prerequisites

Before starting this workflow:

1. **Run the setup script**: `./setup.sh`
2. **Add your data files**:
   - Put Zoom transcript files (`.transcript.vtt`) in `data/transcripts/`
   - Put your roster file (`roster.csv`) in `data/metadata/`
   - Put Zoom recordings list (`zoomus_recordings__*.csv`) in `data/metadata/`
   
   **Note**: The package processes `.transcript.vtt` files for engagement analysis. 
   `.cc.vtt` (closed captions) and `.txt` (chat) files are not processed.

## Step 1: Understanding Your Data

First, let's explore what data you have available:

```{r explore-data}
# Check what transcript files are available
transcript_files <- list.files(
  "data/transcripts",
  pattern = "\\.transcript\\.vtt$",
  full.names = TRUE,
  recursive = TRUE
)

cat("Available transcript files (.transcript.vtt):\n")
basename(transcript_files)

# Also check for other file types (for reference)
cc_files <- list.files(
  "data/transcripts",
  pattern = "\\.cc\\.vtt$",
  full.names = FALSE,
  recursive = TRUE
)

chat_files <- list.files(
  "data/transcripts",
  pattern = "\\.txt$",
  full.names = FALSE,
  recursive = TRUE
)

cat("\nOther file types found:\n")
if (length(cc_files) > 0) {
  cat("- Closed caption files (.cc.vtt):", paste(cc_files, collapse = ", "), "\n")
}
if (length(chat_files) > 0) {
  cat("- Chat files (.txt):", paste(chat_files, collapse = ", "), "\n")
}
cat("\nNote: Only .transcript.vtt files are processed for engagement analysis.\n")

# Check what metadata files are available
cat("\nAvailable metadata files:\n")
list.files("data/metadata")

# Check for specific file types
roster_files <- list.files("data/metadata", pattern = "roster\\.csv$", full.names = TRUE)
zoom_files <- list.files("data/metadata", pattern = "zoomus_recordings__.*\\.csv$", full.names = TRUE)

cat("\nRoster files found:", length(roster_files), "\n")
cat("Zoom recordings files found:", length(zoom_files), "\n")
```

## Step 2: Loading Your Student Roster

Start by loading your student roster:

```{r load-roster}
# Load the student roster
roster_file <- file.path("data/metadata", "roster.csv")

if (file.exists(roster_file)) {
  roster <- load_roster(roster_file)
  
  # View the roster structure
  head(roster)
  cat("\nRoster dimensions:", nrow(roster), "students\n")
  cat("Roster columns:", paste(names(roster), collapse = ", "), "\n")
} else {
  cat("⚠️  Roster file not found at:", roster_file, "\n")
  cat("Please add your roster.csv file to data/metadata/\n")
}
```

## Step 3: Loading and Processing Transcript Files

Now let's load and process your Zoom transcripts:

```{r load-transcripts}
# Find a transcript file to work with
transcript_files <- list.files(
  "data/transcripts",
  pattern = "\\.transcript\\.vtt$",
  full.names = TRUE,
  recursive = TRUE
)

if (length(transcript_files) > 0) {
  # Use the first transcript file
  transcript_file <- transcript_files[1]
  cat("Processing transcript:", basename(transcript_file), "\n")
  
  # Load and process the transcript
  transcript_data <- load_zoom_transcript(transcript_file)
  
  # View the structure
  head(transcript_data)
  cat("\nTranscript dimensions:", nrow(transcript_data), "utterances\n")
  cat("Transcript columns:", paste(names(transcript_data), collapse = ", "), "\n")
} else {
  cat("⚠️  No transcript files found in data/transcripts/\n")
  cat("Please add your Zoom transcript files to data/transcripts/\n")
}
```

## Step 4: Processing Multiple Sessions

If you have multiple sessions, you can process them all at once:

```{r process-multiple}
# Get all transcript files
transcript_files <- list.files(
  "data/transcripts",
  pattern = "\\.transcript\\.vtt$",
  full.names = TRUE,
  recursive = TRUE
)

if (length(transcript_files) > 1) {
  cat("Processing", length(transcript_files), "transcript files...\n")
  
  # Process all transcripts
  all_transcripts <- lapply(transcript_files, function(file) {
    cat("Processing:", basename(file), "\n")
    tryCatch({
      load_zoom_transcript(file)
    }, error = function(e) {
      cat("Error processing", basename(file), ":", e$message, "\n")
      NULL
    })
  })
  
  # Remove any failed transcript loads
  all_transcripts <- all_transcripts[!sapply(all_transcripts, is.null)]
  
  if (length(all_transcripts) > 0) {
    # Combine all transcripts
    combined_transcripts <- dplyr::bind_rows(all_transcripts, .id = "session")
    
    cat("\nCombined data dimensions:", nrow(combined_transcripts), "utterances\n")
    cat("Sessions processed:", length(all_transcripts), "\n")
  } else {
    cat("⚠️  No transcripts were successfully processed\n")
  }
} else {
  cat("Only one transcript file found. Skipping batch processing.\n")
}
```

## Step 5: Calculating Engagement Metrics

Now let's calculate engagement metrics for each student:

```{r calculate-metrics}
# Calculate metrics for all transcript files
transcript_files <- list.files(
  "data/transcripts",
  pattern = "\\.transcript\\.vtt$",
  full.names = FALSE,
  recursive = TRUE
)

if (length(transcript_files) > 0) {
  cat("Calculating metrics for", length(transcript_files), "transcript files...\n")
  
  metrics <- summarize_transcript_files(
    transcript_file_names = transcript_files,
    data_folder = "data",
    transcripts_folder = "transcripts",
    names_to_exclude = c("dead_air", "Unknown")
  )
  
  # View the metrics
  head(metrics)
  cat("\nMetrics dimensions:", nrow(metrics), "participants\n")
  cat("Metrics columns:", paste(names(metrics), collapse = ", "), "\n")
} else {
  cat("⚠️  No transcript files found for metrics calculation\n")
}
```

## Step 6: Matching Student Names

This is often the most challenging part - matching Zoom names to your roster:

```{r name-matching}
if (exists("metrics") && nrow(metrics) > 0 && exists("roster")) {
  # Get names from transcripts
  transcript_names <- unique(metrics$name)
  cat("Names in transcripts:\n")
  print(transcript_names)
  
  # Get names from roster
  roster_names <- unique(roster$preferred_name)
  cat("\nNames in roster:\n")
  print(roster_names)
  
  # Find exact matches
  exact_matches <- intersect(transcript_names, roster_names)
  cat("\nExact matches found:", length(exact_matches), "\n")
  print(exact_matches)
  
  # Find unmatched names
  unmatched_transcript <- setdiff(transcript_names, roster_names)
  unmatched_roster <- setdiff(roster_names, transcript_names)
  
  cat("\nNames in transcripts but not in roster:\n")
  print(unmatched_transcript)
  
  cat("\nNames in roster but not in transcripts:\n")
  print(unmatched_roster)
} else {
  cat("⚠️  Cannot perform name matching - missing metrics or roster data\n")
}
```

## Step 7: Creating Name Mappings

For names that don't match exactly, you'll need to create mappings:

```{r create-mappings}
if (exists("metrics") && nrow(metrics) > 0) {
  # Create a template for name mappings
  transcript_names <- unique(metrics$name)
  
  # Create initial mapping (exact matches)
  name_mapping <- data.frame(
    name_to_clean = transcript_names,
    clean_name = transcript_names,  # Start with exact matches
    notes = "",
    stringsAsFactors = FALSE
  )
  
  # If roster exists, update with exact matches
  if (exists("roster")) {
    roster_names <- unique(roster$preferred_name)
    exact_matches <- intersect(transcript_names, roster_names)
    
    name_mapping$clean_name[match(exact_matches, name_mapping$name_to_clean)] <- exact_matches
    name_mapping$notes[match(exact_matches, name_mapping$name_to_clean)] <- "Exact match"
  }
  
  # Show mapping template
  cat("Name mapping template (edit manually as needed):\n")
  print(name_mapping)
  
  # Save mapping template
  write.csv(name_mapping, "name_mappings_template.csv", row.names = FALSE)
  cat("\nName mapping template saved to: name_mappings_template.csv\n")
  cat("Edit this file manually to add your name mappings, then reload it.\n")
} else {
  cat("⚠️  Cannot create name mappings - missing metrics data\n")
}
```

## Step 8: Loading Custom Name Mappings

If you've created custom name mappings, load them here:

```{r load-mappings}
# Check if custom mappings exist
mapping_file <- "name_mappings_template.csv"

if (file.exists(mapping_file)) {
  custom_mappings <- read.csv(mapping_file, stringsAsFactors = FALSE)
  cat("Loaded custom name mappings:\n")
  print(custom_mappings)
  
  # Apply mappings to metrics
  if (exists("metrics") && nrow(metrics) > 0) {
    clean_metrics <- metrics %>%
      left_join(custom_mappings, by = c("name" = "name_to_clean")) %>%
      mutate(
        student_name = coalesce(clean_name, name),
        is_matched = !is.na(clean_name) & clean_name != name
      )
    
    # Summary of matching success
    matching_summary <- clean_metrics %>%
      group_by(is_matched) %>%
      summarise(
        count = n(),
        percentage = n() / nrow(clean_metrics) * 100
      )
    
    cat("\nName matching summary:\n")
    print(matching_summary)
  }
} else {
  cat("No custom name mappings found. Using original names.\n")
  if (exists("metrics")) {
    clean_metrics <- metrics %>%
      mutate(
        student_name = name,
        is_matched = FALSE
      )
  }
}
```

## Step 9: Analyzing Participation Patterns

Now let's analyze participation patterns to understand engagement:

```{r analyze-patterns}
if (exists("clean_metrics") && nrow(clean_metrics) > 0) {
  # Create a summary by student
  student_summary <- clean_metrics %>%
    group_by(student_name) %>%
    summarise(
      total_utterances = sum(n, na.rm = TRUE),
      total_duration = sum(duration, na.rm = TRUE),
      total_words = sum(wordcount, na.rm = TRUE),
      avg_words_per_minute = mean(wpm, na.rm = TRUE),
      participation_rate = total_utterances / nrow(clean_metrics) * 100
    ) %>%
    arrange(desc(total_utterances))
  
  # View the summary
  head(student_summary, 10)
  
  # Basic statistics
  cat("\nParticipation Statistics:\n")
  participation_stats <- student_summary %>%
    summarise(
      mean_utterances = mean(total_utterances, na.rm = TRUE),
      median_utterances = median(total_utterances, na.rm = TRUE),
      sd_utterances = sd(total_utterances, na.rm = TRUE),
      min_utterances = min(total_utterances, na.rm = TRUE),
      max_utterances = max(total_utterances, na.rm = TRUE)
    )
  print(participation_stats)
} else {
  cat("⚠️  Cannot analyze participation patterns - missing clean metrics data\n")
}
```

## Step 10: Visualizing Participation

Create visualizations to understand participation patterns:

```{r visualize-participation}
if (exists("student_summary") && nrow(student_summary) > 0) {
  # Plot participation by utterance count
  p1 <- ggplot(student_summary, aes(x = reorder(student_name, total_utterances), y = total_utterances)) +
    geom_bar(stat = "identity", fill = "steelblue") +
    coord_flip() +
    labs(
      title = "Student Participation by Utterance Count",
      x = "Student Name",
      y = "Number of Utterances"
    ) +
    theme_minimal()
  
  print(p1)
  
  # Plot participation by duration
  p2 <- ggplot(student_summary, aes(x = reorder(student_name, total_duration), y = total_duration / 60)) +
    geom_bar(stat = "identity", fill = "darkgreen") +
    coord_flip() +
    labs(
      title = "Student Participation by Speaking Time",
      x = "Student Name",
      y = "Speaking Time (minutes)"
    ) +
    theme_minimal()
  
  print(p2)
  
  # Save plots
  ggsave("participation_by_utterances.png", p1, width = 10, height = 8)
  ggsave("participation_by_duration.png", p2, width = 10, height = 8)
  
  cat("Plots saved as:\n")
  cat("- participation_by_utterances.png\n")
  cat("- participation_by_duration.png\n")
} else {
  cat("⚠️  Cannot create visualizations - missing student summary data\n")
}
```

## Step 11: Identifying Participation Gaps

Look for students who might need encouragement to participate:

```{r identify-gaps}
if (exists("student_summary") && nrow(student_summary) > 0) {
  # Identify students with low participation
  median_utterances <- median(student_summary$total_utterances, na.rm = TRUE)
  low_participation <- student_summary %>%
    filter(total_utterances < median_utterances) %>%
    arrange(total_utterances)
  
  cat("Students with below-median participation:\n")
  print(low_participation[, c("student_name", "total_utterances", "total_duration")])
  
  # Create participation categories
  participation_categories <- student_summary %>%
    mutate(
      participation_level = case_when(
        total_utterances >= quantile(total_utterances, 0.75, na.rm = TRUE) ~ "High",
        total_utterances >= quantile(total_utterances, 0.25, na.rm = TRUE) ~ "Medium",
        TRUE ~ "Low"
      )
    )
  
  # Summary by participation level
  level_summary <- participation_categories %>%
    group_by(participation_level) %>%
    summarise(
      count = n(),
      percentage = n() / nrow(participation_categories) * 100
    )
  
  cat("\nParticipation Level Distribution:\n")
  print(level_summary)
} else {
  cat("⚠️  Cannot identify participation gaps - missing student summary data\n")
}
```

## Step 12: Creating Actionable Insights

Generate insights that can help improve equitable participation:

```{r actionable-insights}
cat("=== RECOMMENDATIONS FOR EQUITABLE PARTICIPATION ===\n")
cat("1. Students with low participation may benefit from:\n")
cat("   - Direct invitations to contribute\n")
cat("   - Smaller group discussions\n")
cat("   - Alternative participation methods (chat, polls)\n\n")

cat("2. Consider implementing:\n")
cat("   - Structured discussion protocols\n")
cat("   - Think-pair-share activities\n")
cat("   - Anonymous participation options\n\n")

cat("3. Monitor participation patterns over time to:\n")
cat("   - Track improvement in engagement\n")
cat("   - Identify effective interventions\n")
cat("   - Ensure all students feel included\n\n")

if (exists("student_summary") && nrow(student_summary) > 0) {
  cat("4. Specific recommendations for your class:\n")
  cat("   - Total students analyzed:", nrow(student_summary), "\n")
  cat("   - Average utterances per student:", round(mean(student_summary$total_utterances, na.rm = TRUE), 1), "\n")
  cat("   - Students with below-average participation:", sum(student_summary$total_utterances < mean(student_summary$total_utterances, na.rm = TRUE)), "\n")
}
```

## Step 13: Saving Your Analysis

Save your processed data for future reference:

```{r save-analysis}
# Create output directory
if (!dir.exists("outputs")) {
  dir.create("outputs")
}

# Save the clean metrics
if (exists("clean_metrics")) {
  write_engagement_metrics(clean_metrics, "outputs/clean_engagement_metrics.csv", comments_format = "text")
  cat("Saved: outputs/clean_engagement_metrics.csv\n")
}

# Save the student summary
if (exists("student_summary")) {
  write_engagement_metrics(student_summary, "outputs/student_participation_summary.csv")
  cat("Saved: outputs/student_participation_summary.csv\n")
}

# Save the name mappings
if (exists("custom_mappings")) {
  write_engagement_metrics(custom_mappings, "outputs/name_mappings.csv")
  cat("Saved: outputs/name_mappings.csv\n")
}

cat("\nAnalysis files saved to outputs/ directory\n")
```

## Best Practices and Tips

### Data Organization
- Keep your transcript files organized by date/session
- Maintain a consistent naming convention
- Back up your original data files
- Document any manual name mappings

### Ethical Considerations
- Use this data to promote equitable participation, not surveillance
- Focus on group patterns rather than individual performance
- Respect student privacy and preferences
- Share insights constructively with students when appropriate

### Common Pitfalls to Avoid
- Don't assume low participation means disengagement
- Consider cultural and personal communication preferences
- Remember that quality of participation matters more than quantity
- Avoid making assumptions about student abilities based on participation

### Interpreting Results
- Look for patterns, not individual judgments
- Consider context (class size, format, topic)
- Use data to inform teaching strategies, not student evaluation
- Focus on creating inclusive learning environments

## Next Steps

This analysis provides a foundation for understanding participation patterns. Consider:

1. **Regular monitoring**: Track participation patterns over the semester
2. **Intervention strategies**: Implement targeted approaches for low-participation students
3. **Student feedback**: Ask students about their participation preferences
4. **Pedagogical adjustments**: Modify teaching strategies based on insights
5. **Continuous improvement**: Refine your analysis approach over time

Remember: The goal is to create an inclusive learning environment where all students feel comfortable and encouraged to participate in ways that work for them. 