---
title: "Whole Game: Real-World Workflow for Instructors"
author: "zoomstudentengagement package"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document:
    toc: true
    toc_float: true
    code_folding: hide
    theme: cosmo
    highlight: tango
    df_print: paged
    encoding: UTF-8
params:
  instructor_name: "Conor Healy"
---

```{r setup, include = FALSE}
# Set encoding for R Markdown
Sys.setlocale("LC_ALL", "en_US.UTF-8")
options(encoding = "UTF-8")

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6,
  warning = FALSE,
  message = FALSE,
  echo = TRUE
)
```

```{r load-packages}
library(zoomstudentengagement)
library(dplyr)
library(ggplot2)
library(readr)
library(tibble)
library(lubridate)
library(jsonlite)
```

# Whole Game: Real-World Workflow for Instructors

This document demonstrates the complete workflow for analyzing student engagement in Zoom sessions using your actual data. It's designed to work with the real-world testing environment.

## Required Inputs

Before running this workflow, ensure you have the following:

### 1. Instructor Information (YAML Parameters)
- **instructor_name**: Name of the instructor as it appears in Zoom transcripts (e.g., "Conor Healy", "Dr. Healy", "Professor Healy", "Instructor")

### 2. Student Roster (`data/metadata/roster.csv`)
- CSV file with enrolled student information
- Columns: `preferred_name`, `formal_name`, `student_id`
- **Note**: Instructors should NOT be in the roster

### 3. Zoom Transcript Files (`data/transcripts/`)
- `.transcript.vtt` files from Zoom recordings
- Only transcript files are processed (not `.cc.vtt` or `.txt` files)

### 4. Name Lookup Configuration (`data/metadata/section_names_lookup.csv`)
- **Auto-created** from instructor parameters
- Maps Zoom transcript names to preferred names
- Identifies instructors vs. students via `participant_type` column

## How to Run

```bash
# From the scripts/real_world_testing directory:
Rscript -e "rmarkdown::render('whole_game_real_world.Rmd', 
  params = list(instructor_name = 'Instructor Name'))"
```

Or edit the YAML parameters at the top of this file and run:
```bash
Rscript -e "rmarkdown::render('whole_game_real_world.Rmd')"
```

## Prerequisites

Before starting this workflow:

1. **Run the setup script**: `./setup.sh`
2. **Add your data files**:
   - Put Zoom transcript files (`.transcript.vtt`) in `data/transcripts/`
   - Put your roster file (`roster.csv`) in `data/metadata/`
   - Put Zoom recordings list (`zoomus_recordings__*.csv`) in `data/metadata/`
   
   **Note**: The package processes `.transcript.vtt` files for engagement analysis. 
   `.cc.vtt` (closed captions) and `.txt` (chat) files are not processed.

## Step 1: Understanding Your Data

First, let's explore what data you have available:

## 🔒 **Privacy Testing Setup**

Before we begin the main workflow, let's test the privacy features to ensure they're working correctly. This is crucial for FERPA compliance and ethical use.

### **Testing Privacy Levels**

The package supports four privacy levels that control how names are handled:

1. **`ferpa_strict`**: Maximum privacy - masks all names including instructors
2. **`ferpa_standard`**: Standard privacy - masks student names, preserves instructor names
3. **`mask`**: Basic masking - masks student names, preserves instructor names
4. **`none`**: No masking - shows all names (use with extreme caution)

Let's test each privacy level:

```{r explore-data}
# Check what transcript files are available
transcript_files <- list.files(
  "data/transcripts",
  pattern = "\\.transcript\\.vtt$",
  full.names = TRUE,
  recursive = TRUE
)

cat("Available transcript files (.transcript.vtt):\n")
basename(transcript_files)

# Also check for other file types (for reference)
cc_files <- list.files(
  "data/transcripts",
  pattern = "\\.cc\\.vtt$",
  full.names = FALSE,
  recursive = TRUE
)

chat_files <- list.files(
  "data/transcripts",
  pattern = "\\.txt$",
  full.names = FALSE,
  recursive = TRUE
)

cat("\nOther file types found:\n")
if (length(cc_files) > 0) {
  cat("- Closed caption files (.cc.vtt):", paste(cc_files, collapse = ", "), "\n")
}
if (length(chat_files) > 0) {
  cat("- Chat files (.txt):", paste(chat_files, collapse = ", "), "\n")
}
cat("\nNote: Only .transcript.vtt files are processed for engagement analysis.\n")

# Check what metadata files are available
cat("\nAvailable metadata files:\n")
list.files("data/metadata")

# Check for specific file types
roster_files <- list.files("data/metadata", pattern = "roster\\.csv$", full.names = TRUE)
zoom_files <- list.files("data/metadata", pattern = "zoomus_recordings__.*\\.csv$", full.names = TRUE)

cat("\nRoster files found:", length(roster_files), "\n")
cat("Zoom recordings files found:", length(zoom_files), "\n")
```

### **Step 1.1: Privacy Level Testing**

Now let's test each privacy level to ensure they work correctly:

```{r test-privacy-levels}
# Function to test privacy levels
test_privacy_level <- function(level, description) {
  cat(sprintf("\n=== Testing %s (%s) ===\n", level, description))
  
      # Set privacy level
    set_privacy_defaults(level)
    
    # Verify privacy level was set
    current_level <- getOption("zoomstudentengagement.privacy_level", "mask")
    cat("Privacy level set to:", current_level, "\n")
    
    # Load a transcript file if available
    transcript_files <- list.files(
      "data/transcripts",
      pattern = "\\.transcript\\.vtt$",
      full.names = TRUE,
      recursive = TRUE
    )
    
    if (length(transcript_files) > 0) {
      transcript_file <- transcript_files[1]
      cat("Testing with:", basename(transcript_file), "\n")
      
      # Calculate metrics with current privacy level
      metrics <- summarize_transcript_metrics(
        transcript_file_path = transcript_file,
        names_exclude = c("dead_air")
      )
    
    # Check what names we see
    unique_names <- unique(metrics$name)
    cat("Unique names found:", length(unique_names), "\n")
    cat("Sample names:", paste(head(unique_names, 5), collapse = ", "), "\n")
    
    # Debug: Show ALL names to see what we're missing
    cat("All unique names:\n")
    name_df <- data.frame(
      Index = seq_along(unique_names),
      Name = unique_names,
      stringsAsFactors = FALSE
    )
    print(name_df)
    
    # Check for real names vs masked names
    real_name_pattern <- "^[A-Z][a-z]+(\\s+[A-Z][a-z]+)*$"
    masked_pattern <- "^Student\\s+\\d{2}$"
    
    real_names <- unique_names[grepl(real_name_pattern, unique_names)]
    masked_names <- unique_names[grepl(masked_pattern, unique_names)]
    
    cat("Real names found:", length(real_names), "\n")
    cat("Masked names found:", length(masked_names), "\n")
    
    # Debug: Show which names match each pattern
    cat("Names matching real name pattern:\n")
    for (name in real_names) {
      cat(sprintf("  '%s'\n", name))
    }
    
    cat("Names matching masked pattern:\n")
    for (name in masked_names) {
      cat(sprintf("  '%s'\n", name))
    }
    
    # Show names that don't match either pattern
    unmatched_names <- unique_names[!grepl(real_name_pattern, unique_names) & !grepl(masked_pattern, unique_names)]
    if (length(unmatched_names) > 0) {
      cat("Names that don't match either pattern:\n")
      for (name in unmatched_names) {
        cat(sprintf("  '%s'\n", name))
      }
    }
    
    # Additional check: Test privacy function directly
    cat("\n--- Direct Privacy Function Test ---\n")
    # Create a test dataframe with real names
    test_df <- data.frame(
      name = c("John Smith", "Jane Doe", "Bob Wilson"),
      value = c(1, 2, 3),
      stringsAsFactors = FALSE
    )
    
    # Test privacy function directly
    masked_df <- ensure_privacy(test_df, privacy_level = level)
    direct_names <- unique(masked_df$name)
    cat("Direct privacy test - Sample names:", paste(direct_names, collapse = ", "), "\n")
    
    direct_real <- direct_names[grepl(real_name_pattern, direct_names)]
    direct_masked <- direct_names[grepl(masked_pattern, direct_names)]
    cat("Direct test - Real names:", length(direct_real), "Masked names:", length(direct_masked), "\n")
    
    # Validate expected behavior
    if (level == "ferpa_strict" || level == "ferpa_standard") {
      if (length(real_names) > 0) {
        cat("⚠️  WARNING: Real names found in", level, "mode\n")
      } else {
        cat("✅ Privacy level", level, "working correctly\n")
      }
    } else if (level == "mask") {
      if (length(masked_names) > 0) {
        cat("✅ Privacy level", level, "working correctly\n")
      } else {
        cat("⚠️  WARNING: No masked names found in", level, "mode\n")
      }
    } else if (level == "none") {
      if (length(real_names) > 0) {
        cat("✅ Privacy level", level, "working correctly (showing real names)\n")
      } else {
        cat("⚠️  WARNING: No real names found in", level, "mode\n")
      }
    }
    
    return(metrics)
  } else {
    cat("⚠️  No transcript files available for testing\n")
    return(NULL)
  }
}

# Test all privacy levels
cat("🔒 Testing Privacy Levels\n")
cat("========================\n")

ferpa_strict_metrics <- test_privacy_level("ferpa_strict", "Maximum privacy - masks all names")
ferpa_standard_metrics <- test_privacy_level("ferpa_standard", "Standard privacy - masks students, preserves instructors")
mask_metrics <- test_privacy_level("mask", "Basic masking - masks students, preserves instructors")
none_metrics <- test_privacy_level("none", "No masking - shows all names (use with caution)")

# Reset to default privacy level
set_privacy_defaults("mask")
cat("\n✅ Privacy testing completed. Reset to default 'mask' level.\n")
```

### **Step 1.2: FERPA Compliance Validation**

Let's validate that our outputs are FERPA compliant:

```{r ferpa-validation}
# Function to check for PII in outputs
check_ferpa_compliance <- function(data, description) {
  cat(sprintf("\n=== FERPA Compliance Check: %s ===\n", description))
  
  # Convert data to text for scanning
  data_text <- paste(capture.output(print(data)), collapse = " ")
  
  # Check for PII indicators
  pii_indicators <- c("email", "@", "phone", "address", "ssn", "id", "student_id")
  pii_found <- sapply(pii_indicators, function(x) grepl(x, data_text, ignore.case = TRUE))
  
  if (any(pii_found)) {
    cat("🚨 FERPA VIOLATION: PII found in output\n")
    cat("PII indicators found:", paste(names(pii_found[pii_found]), collapse = ", "), "\n")
    return(FALSE)
  } else {
    cat("✅ FERPA compliant: No PII detected\n")
    return(TRUE)
  }
}

# Check FERPA compliance for each privacy level
if (!is.null(ferpa_strict_metrics)) {
  check_ferpa_compliance(ferpa_strict_metrics, "ferpa_strict level")
}

if (!is.null(ferpa_standard_metrics)) {
  check_ferpa_compliance(ferpa_standard_metrics, "ferpa_standard level")
}

if (!is.null(mask_metrics)) {
  check_ferpa_compliance(mask_metrics, "mask level")
}

# Test export security
cat("\n=== Testing Export Security ===\n")
if (!is.null(mask_metrics)) {
  # Test that exported files don't contain real names
  export_file <- "test_privacy_export.csv"
  write.csv(mask_metrics, export_file, row.names = FALSE)
  
  export_content <- readLines(export_file)
  export_has_real_names <- any(sapply(export_content, function(x) grepl("^[A-Z][a-z]+ [A-Z][a-z]+", x)))
  
  if (export_has_real_names) {
    cat("🚨 SECURITY ISSUE: Real names found in exported file\n")
  } else {
    cat("✅ Export security: No real names in exported file\n")
  }
  
  # Clean up test file
  unlink(export_file)
}
```

## Step 2: Configuring Instructor Names (Recommended First Step)

**Important**: Before running the main workflow, it's recommended to configure instructor names in the `section_names_lookup.csv` file. This prevents expected exceptions and makes the process smoother.

```{r configure-instructors}
# Function to create instructor entries
create_instructor_entries <- function(name) {
  # Only use the exact instructor name provided
  # Users can add variations manually after seeing what appears in transcripts
  data.frame(
    transcript_name = name,
    preferred_name = name,
    formal_name = name,
    participant_type = "instructor",
    student_id = "INSTRUCTOR",
    notes = "Primary instructor name - add variations manually if needed",
    stringsAsFactors = FALSE
  )
}

# Function to safely update lookup file with backup
safe_update_lookup_file <- function(new_data, file_path) {
  # Create backup if file exists
  if (file.exists(file_path)) {
    backup_path <- paste0(file_path, ".backup.", format(Sys.time(), "%Y%m%d_%H%M%S"))
    file.copy(file_path, backup_path)
    cat("📁 Backup created:", basename(backup_path), "\n")
  }
  
  # Write new data
  write.csv(new_data, file_path, row.names = FALSE)
}

# Get instructor information from YAML parameters
instructor_name <- params$instructor_name

cat("=== Instructor Configuration ===\n")
cat("Instructor Name:", instructor_name, "\n")

# Check if section_names_lookup.csv exists
lookup_file <- file.path("data/metadata", "section_names_lookup.csv")

if (file.exists(lookup_file)) {
  cat("\n=== Existing Name Lookup Configuration ===\n")
  existing_lookup <- read.csv(lookup_file, stringsAsFactors = FALSE)
  cat("Found existing lookup file with", nrow(existing_lookup), "entries\n")
  
  # Check for instructor entries
  if ("participant_type" %in% names(existing_lookup)) {
    instructor_entries <- existing_lookup[existing_lookup$participant_type == "instructor", ]
    if (nrow(instructor_entries) > 0) {
      cat("✅ Instructors already configured:\n")
      knitr::kable(instructor_entries[, c("transcript_name", "preferred_name", "participant_type")], 
                  caption = "Configured Instructors")
      
      # Check if current instructor is already configured
      current_instructor_configured <- any(instructor_entries$preferred_name == instructor_name)
      if (!current_instructor_configured) {
        cat("⚠️  Current instructor not found in existing configuration.\n")
        cat("Adding current instructor to configuration...\n")
        
        # Add current instructor to existing configuration (preserve existing entries)
        new_instructor_entries <- create_instructor_entries(instructor_name)
        updated_lookup <- rbind(existing_lookup, new_instructor_entries)
        safe_update_lookup_file(updated_lookup, lookup_file)
        cat("✅ Updated instructor configuration (preserved existing entries).\n")
      }
    } else {
      cat("⚠️  No instructors configured in lookup file\n")
      cat("Adding current instructor to configuration (preserving existing entries)...\n")
      
      # Add instructor entries to existing configuration (don't overwrite)
      instructor_lookup <- create_instructor_entries(instructor_name)
      updated_lookup <- rbind(existing_lookup, instructor_lookup)
      safe_update_lookup_file(updated_lookup, lookup_file)
      cat("✅ Instructor configuration added (preserved existing entries).\n")
    }
  } else {
    cat("⚠️  Lookup file missing participant_type column.\n")
    
    # Check if the file is empty
    if (nrow(existing_lookup) == 0) {
      cat("File is empty. Creating new lookup file with instructor configuration...\n")
      instructor_lookup <- create_instructor_entries(instructor_name)
      safe_update_lookup_file(instructor_lookup, lookup_file)
      cat("✅ Created new lookup file with instructor configuration.\n")
    } else {
      cat("Adding participant_type column and instructor configuration (preserving existing entries)...\n")
      
      # Add participant_type column to existing entries and add instructor
      existing_lookup$participant_type <- "enrolled_student"  # Default for existing entries
      instructor_lookup <- create_instructor_entries(instructor_name)
      updated_lookup <- rbind(existing_lookup, instructor_lookup)
      safe_update_lookup_file(updated_lookup, lookup_file)
      cat("✅ Added participant_type column and instructor configuration (preserved existing entries).\n")
    }
  }
} else {
  cat("\n=== Creating Instructor Configuration ===\n")
  cat("No section_names_lookup.csv found. Creating one with instructor configuration.\n")
  
  # Create instructor entries and save
  instructor_lookup <- create_instructor_entries(instructor_name)
  safe_update_lookup_file(instructor_lookup, lookup_file)
  cat("✅ Instructor configuration created:", lookup_file, "\n")
cat("This file maps Zoom transcript names to your instructor's preferred name.\n")
cat("Edit this file to add more instructor name variations or other participants.\n")
cat("💡 Note: Automatic backups are created when the file is updated.\n")
}

cat("\n=== Instructor Configuration Instructions ===\n")
cat("The section_names_lookup.csv file contains:\n")
cat("- transcript_name: Names as they appear in Zoom transcripts\n")
cat("- preferred_name: How you want them displayed in reports\n")
cat("- participant_type: 'instructor' for instructors, 'enrolled_student' for students\n")
cat("- student_id: Student ID or 'INSTRUCTOR' for instructors\n")
cat("- notes: Any additional information\n")
cat("\nCurrent instructor configuration:\n")
if (exists("instructor_lookup")) {
  knitr::kable(instructor_lookup[, c("transcript_name", "preferred_name", "participant_type")], 
              caption = "Current Instructor Configuration")
} else if (file.exists(lookup_file)) {
  current_lookup <- read.csv(lookup_file, stringsAsFactors = FALSE)
  instructor_entries <- current_lookup[current_lookup$participant_type == "instructor", ]
  if (nrow(instructor_entries) > 0) {
    knitr::kable(instructor_entries[, c("transcript_name", "preferred_name", "participant_type")], 
                caption = "Current Instructor Configuration")
  }
}

cat("\n=== Adding Instructor Name Variations ===\n")
cat("After processing transcripts, you may need to add variations of the instructor name.\n")
cat("Common reasons to add variations:\n")
cat("- Zoom shows 'Dr. Healy' instead of 'Conor Healy'\n")
cat("- Zoom shows 'Instructor' or 'Host' for the instructor\n")
cat("- Different sessions show different name formats\n")
cat("\nTo add variations, edit data/metadata/section_names_lookup.csv and add new rows with:\n")
cat("- transcript_name: The name as it appears in Zoom\n")
cat("- preferred_name: The instructor's preferred name\n")
cat("- participant_type: 'instructor'\n")
```

## Step 3: Loading Your Student Roster

Start by loading your student roster:

```{r load-roster}
# Load the student roster
roster_file <- "roster.csv"
roster_path <- file.path("data/metadata", roster_file)

if (file.exists(roster_path)) {
  tryCatch({
    roster <- load_roster(data_folder = "data/metadata", roster_file = roster_file)
    
    # View the roster structure
    knitr::kable(head(roster), caption = "Roster Preview")
    cat("\nRoster dimensions:", nrow(roster), "students\n")
    cat("Roster columns:", paste(names(roster), collapse = ", "), "\n")
    
    # Note about instructor configuration
    cat("\n=== Note: Instructor Configuration ===\n")
    cat("Instructors should be configured in section_names_lookup.csv (Step 2),\n")
    cat("not in the roster.csv file. The roster is for enrolled students only.\n")
    
  }, error = function(e) {
    cat("⚠️  Error loading roster:", e$message, "\n")
    cat("Please check that your roster.csv file is properly formatted\n")
  })
} else {
  cat("⚠️  Roster file not found at:", roster_path, "\n")
  cat("Please add your roster.csv file to data/metadata/\n")
  
  # Create a sample roster template (students only)
  cat("\n=== Creating Sample Roster Template ===\n")
  sample_roster <- data.frame(
    preferred_name = c("Student One", "Student Two", "Student Three"),
    formal_name = c("Student One", "Student Two", "Student Three"),
    student_id = c("STU001", "STU002", "STU003"),
    stringsAsFactors = FALSE
  )
  
  # Save sample roster
  sample_roster_path <- file.path("data/metadata", "sample_roster.csv")
  write.csv(sample_roster, sample_roster_path, row.names = FALSE)
  cat("Sample roster template created:", sample_roster_path, "\n")
  cat("Edit this file to match your actual student roster.\n")
  cat("Note: Instructors should be configured in section_names_lookup.csv (Step 2).\n")
}
```

## Step 3: Loading and Processing Transcript Files

Now let's load and process your Zoom transcripts:

```{r load-transcripts}
# Find a transcript file to work with
transcript_files <- list.files(
  "data/transcripts",
  pattern = "\\.transcript\\.vtt$",
  full.names = TRUE,
  recursive = TRUE
)

if (length(transcript_files) > 0) {
  # Use the first transcript file
  transcript_file <- transcript_files[1]
  cat("Processing transcript:", basename(transcript_file), "\n")
  
  # Load and process the transcript
  transcript_data <- load_zoom_transcript(transcript_file)
  
      # View the structure
    knitr::kable(head(transcript_data), caption = "Transcript Preview")
    cat("\nTranscript dimensions:", nrow(transcript_data), "utterances\n")
    cat("Transcript columns:", paste(names(transcript_data), collapse = ", "), "\n")
} else {
  cat("⚠️  No transcript files found in data/transcripts/\n")
  cat("Please add your Zoom transcript files to data/transcripts/\n")
}
```

## Step 4: Processing Multiple Sessions

If you have multiple sessions, you can process them all at once:

```{r process-multiple}
# Get all transcript files
transcript_files <- list.files(
  "data/transcripts",
  pattern = "\\.transcript\\.vtt$",
  full.names = TRUE,
  recursive = TRUE
)

if (length(transcript_files) > 1) {
  cat("Processing", length(transcript_files), "transcript files...\n")
  
  # Process all transcripts
  all_transcripts <- lapply(transcript_files, function(file) {
    cat("Processing:", basename(file), "\n")
    tryCatch({
      load_zoom_transcript(file)
    }, error = function(e) {
      cat("Error processing", basename(file), ":", e$message, "\n")
      NULL
    })
  })
  
  # Remove any failed transcript loads
  all_transcripts <- all_transcripts[!sapply(all_transcripts, is.null)]
  
  if (length(all_transcripts) > 0) {
    # Combine all transcripts
    combined_transcripts <- dplyr::bind_rows(all_transcripts, .id = "session")
    
    cat("\nCombined data dimensions:", nrow(combined_transcripts), "utterances\n")
    cat("Sessions processed:", length(all_transcripts), "\n")
  } else {
    cat("⚠️  No transcripts were successfully processed\n")
  }
} else {
  cat("Only one transcript file found. Skipping batch processing.\n")
}
```

## Step 5: Calculating Engagement Metrics

Now let's calculate engagement metrics for each student:

```{r calculate-metrics}
# Calculate metrics for all transcript files
transcript_files <- list.files(
  "data/transcripts",
  pattern = "\\.transcript\\.vtt$",
  full.names = FALSE,
  recursive = TRUE
)

if (length(transcript_files) > 0) {
  cat("Calculating metrics for", length(transcript_files), "transcript files...\n")
  
  metrics <- summarize_transcript_files(
    transcript_file_names = transcript_files,
    data_folder = "data",
    transcripts_folder = "transcripts",
    names_to_exclude = c("dead_air", "Unknown")
  )
  
  # View the metrics
  knitr::kable(head(metrics), caption = "Metrics Preview")
  cat("\nMetrics dimensions:", nrow(metrics), "participants\n")
  cat("Metrics columns:", paste(names(metrics), collapse = ", "), "\n")
} else {
  cat("⚠️  No transcript files found for metrics calculation\n")
}
```

## Step 6: Instructor Identification and Name Matching

This is a critical step that combines instructor identification with privacy-aware name matching. The package now includes sophisticated name matching capabilities that maintain privacy while helping you match Zoom names to your roster.

### **Step 6.1: Instructor Configuration Check**

Let's verify that instructors are properly configured in the lookup file:

```{r instructor-configuration-check}
# Check instructor configuration in section_names_lookup.csv
lookup_file <- file.path("data/metadata", "section_names_lookup.csv")

if (file.exists(lookup_file)) {
  cat("=== Instructor Configuration Check ===\n")
  lookup_data <- read.csv(lookup_file, stringsAsFactors = FALSE)
  
  if ("participant_type" %in% names(lookup_data)) {
    instructor_entries <- lookup_data[lookup_data$participant_type == "instructor", ]
    
    if (nrow(instructor_entries) > 0) {
      cat("✅ Instructors configured in lookup file:\n")
      print(instructor_entries[, c("transcript_name", "preferred_name", "participant_type")])
      
      # Check if any of these names appear in the transcript
      if (exists("transcript_data")) {
        transcript_speakers <- unique(transcript_data$name)
        configured_instructor_names <- instructor_entries$transcript_name
        
        matching_instructors <- intersect(transcript_speakers, configured_instructor_names)
        if (length(matching_instructors) > 0) {
          cat("✅ Found configured instructors in transcript:", paste(matching_instructors, collapse = ", "), "\n")
        } else {
          cat("⚠️  Configured instructors not found in transcript.\n")
          cat("Transcript speakers:", paste(transcript_speakers, collapse = ", "), "\n")
          cat("Consider adding more name variations to section_names_lookup.csv\n")
        }
      }
    } else {
      cat("⚠️  No instructors configured in lookup file.\n")
      cat("Please add instructor entries to section_names_lookup.csv before proceeding.\n")
    }
  } else {
    cat("⚠️  Lookup file missing participant_type column.\n")
    cat("Please ensure section_names_lookup.csv has a participant_type column.\n")
  }
} else {
  cat("⚠️  No section_names_lookup.csv found.\n")
  cat("Please run Step 2 to configure instructor names first.\n")
}
```

### **Step 6.2: Privacy-Aware Name Matching Workflow**

Now let's use the new privacy-aware name matching workflow. This is the recommended approach that maintains privacy while helping you match names:

```{r privacy-aware-name-matching}
# Use the safe name matching workflow for privacy-aware matching
if (exists("transcript_data") && exists("roster")) {
  cat("\n=== Privacy-Aware Name Matching ===\n")
  
  # Get the first transcript file for processing
  transcript_files <- list.files(
    "data/transcripts",
    pattern = "\\.transcript\\.vtt$",
    full.names = TRUE,
    recursive = TRUE
  )
  
  if (length(transcript_files) > 0) {
    transcript_file <- transcript_files[1]
    cat("Processing transcript:", basename(transcript_file), "\n")
    
    # Set up privacy defaults for name matching
    set_privacy_defaults(
      privacy_level = "ferpa_standard",  # Standard privacy for matching
      unmatched_names_action = "warn"    # Warn about unmatched names
    )
    
    # Run the safe name matching workflow
    tryCatch({
      cat("Running privacy-aware name matching workflow...\n")
      
      # First, try with default settings (stops on unmatched names)
      result <- safe_name_matching_workflow(
        transcript_file_path = transcript_file,
        roster_data = roster,
        privacy_level = "ferpa_standard",
        unmatched_names_action = "warn",  # This will warn and guide through matching
        data_folder = "data/metadata"  # Use the same directory as instructor configuration
      )
      
      cat("✅ Name matching workflow completed successfully!\n")
      
      # Show matching results
      if (!is.null(result) && nrow(result) > 0) {
        cat("\nName Matching Results:\n")
        cat("Total participants:", nrow(result), "\n")
        
        # Count by participant type
        if ("participant_type" %in% names(result)) {
          type_counts <- table(result$participant_type)
          cat("Participant types:\n")
          for (type in names(type_counts)) {
            cat("  ", type, ":", type_counts[type], "\n")
          }
        }
        
        # Show matched vs unmatched
        if ("is_matched" %in% names(result)) {
          matched_count <- sum(result$is_matched, na.rm = TRUE)
          unmatched_count <- sum(!result$is_matched, na.rm = TRUE)
          cat("Matched names:", matched_count, "\n")
          cat("Unmatched names:", unmatched_count, "\n")
        }
        
        # Store the result for later use
        clean_metrics <- result
        
      } else {
        cat("⚠️  No results returned from name matching workflow\n")
      }
      
    }, error = function(e) {
      cat("❌ Error in name matching workflow:", e$message, "\n")
      cat("This may indicate unmatched names that need manual resolution.\n")
      
      # Try to provide guidance for manual matching
      cat("\n=== Manual Name Matching Guidance ===\n")
      cat("If you have unmatched names, you can:\n")
      cat("1. Use prompt_name_matching() to create a lookup template\n")
      cat("2. Edit the template manually with correct name mappings\n")
      cat("3. Re-run the workflow with the updated mappings\n")
    })
    
    # Ensure instructor is always in the lookup file
    if (file.exists(lookup_file)) {
      cat("\n=== Ensuring Instructor Configuration ===\n")
      current_lookup <- read.csv(lookup_file, stringsAsFactors = FALSE)
      
      # Check if instructor is already configured
      if ("participant_type" %in% names(current_lookup)) {
        instructor_entries <- current_lookup[current_lookup$participant_type == "instructor", ]
        if (nrow(instructor_entries) == 0) {
          cat("Adding instructor configuration to lookup file (preserving existing entries)...\n")
          instructor_entry <- create_instructor_entries(instructor_name)
          updated_lookup <- rbind(current_lookup, instructor_entry)
          safe_update_lookup_file(updated_lookup, lookup_file)
          cat("✅ Instructor configuration added to lookup file (preserved existing entries).\n")
        } else {
          cat("✅ Instructor already configured in lookup file.\n")
        }
      } else {
        cat("⚠️  Lookup file missing participant_type column.\n")
        
        # Check if the file is empty
        if (nrow(current_lookup) == 0) {
          cat("File is empty. Creating new lookup file with instructor configuration...\n")
          instructor_entry <- create_instructor_entries(instructor_name)
          safe_update_lookup_file(instructor_entry, lookup_file)
          cat("✅ Created new lookup file with instructor configuration.\n")
        } else {
          cat("Adding participant_type and instructor configuration (preserving existing entries)...\n")
          current_lookup$participant_type <- "enrolled_student"  # Default for existing entries
          instructor_entry <- create_instructor_entries(instructor_name)
          updated_lookup <- rbind(current_lookup, instructor_entry)
          safe_update_lookup_file(updated_lookup, lookup_file)
          cat("✅ Added participant_type column and instructor configuration (preserved existing entries).\n")
        }
      }
    }
    
  } else {
    cat("⚠️  No transcript files found for name matching\n")
  }
} else {
  cat("⚠️  Cannot perform name matching - missing transcript data or roster\n")
}
```

### **Step 6.3: Manual Name Matching (if needed)**

**Note**: Manual name matching is handled automatically by the privacy-aware workflow in Step 6.2. If unmatched names are found, the workflow will create a `section_names_lookup.csv` file in the `data/` directory with instructions for manual mapping.

If you need to manually create or update name mappings, you can:

1. **Edit the existing file**: `data/section_names_lookup.csv` (created by the workflow)
2. **Add instructor variations**: If the instructor appears with different names in transcripts
3. **Map student names**: Link transcript names to roster names
4. **Set participant types**: Use 'instructor', 'enrolled_student', or 'guest'

The workflow will automatically detect and use your manual mappings on the next run.

### **Step 6.4: Loading Custom Name Mappings**

If you've created custom name mappings, load them here:

```{r load-custom-mappings}
# Check for custom name mappings
mapping_files <- list.files(
  "data/metadata",
  pattern = ".*names.*lookup.*csv$",
  full.names = TRUE
)

if (length(mapping_files) > 0) {
  cat("\n=== Loading Custom Name Mappings ===\n")
  
  # Use the most recent mapping file
  mapping_file <- mapping_files[length(mapping_files)]
  cat("Loading mappings from:", basename(mapping_file), "\n")
  
  tryCatch({
    custom_mappings <- read.csv(mapping_file, stringsAsFactors = FALSE)
    
    cat("Loaded", nrow(custom_mappings), "name mappings\n")
    
    # Show a preview of the mappings
    if (nrow(custom_mappings) > 0) {
      cat("Mapping preview:\n")
      print(head(custom_mappings))
    }
    
    # Apply mappings to transcript data if available
    if (exists("transcript_data")) {
      clean_transcript <- transcript_data %>%
        left_join(custom_mappings, by = c("name" = "transcript_name")) %>%
        mutate(
          clean_name = coalesce(preferred_name, name),
          is_matched = !is.na(preferred_name) & preferred_name != name
        )
      
      # Summary of matching success
      if ("is_matched" %in% names(clean_transcript)) {
        matched_count <- sum(clean_transcript$is_matched, na.rm = TRUE)
        total_count <- nrow(clean_transcript)
        cat("Applied mappings:", matched_count, "of", total_count, "utterances matched\n")
      }
      
      # Store for later use
      clean_metrics <- clean_transcript
    }
    
  }, error = function(e) {
    cat("❌ Error loading custom mappings:", e$message, "\n")
  })
} else {
  cat("No custom name mapping files found.\n")
}
```

## Step 7: Analyzing Participation Patterns

Now let's analyze participation patterns to understand engagement:

```{r analyze-patterns}
if (exists("clean_metrics") && nrow(clean_metrics) > 0) {
  # Create a summary by student
  # Use the appropriate name column based on what's available
  name_col <- if ("preferred_name" %in% names(clean_metrics)) "preferred_name" else "name"
  
  # Check what columns are available
  available_cols <- names(clean_metrics)
  cat("Available columns in clean_metrics:", paste(available_cols, collapse = ", "), "\n")
  
  # Create a simple summary based on what's available
  if ("duration" %in% available_cols) {
    student_summary <- clean_metrics %>%
      group_by(!!sym(name_col)) %>%
      summarise(
        total_duration = sum(duration, na.rm = TRUE),
        utterance_count = n(),
        participation_rate = n() / nrow(clean_metrics) * 100
      ) %>%
      arrange(desc(total_duration))
  } else {
    # Fallback to just counting utterances
    student_summary <- clean_metrics %>%
      group_by(!!sym(name_col)) %>%
      summarise(
        utterance_count = n(),
        participation_rate = n() / nrow(clean_metrics) * 100
      ) %>%
      arrange(desc(utterance_count))
  }
  
  # View the summary
  knitr::kable(head(student_summary, 10), caption = "Student Participation Summary")
  
  # Basic statistics
  cat("\nParticipation Statistics:\n")
  if ("total_duration" %in% names(student_summary)) {
    participation_stats <- student_summary %>%
      summarise(
        mean_duration = mean(total_duration, na.rm = TRUE),
        median_duration = median(total_duration, na.rm = TRUE),
        mean_utterances = mean(utterance_count, na.rm = TRUE),
        median_utterances = median(utterance_count, na.rm = TRUE)
      )
  } else {
    participation_stats <- student_summary %>%
      summarise(
        mean_utterances = mean(utterance_count, na.rm = TRUE),
        median_utterances = median(utterance_count, na.rm = TRUE)
      )
  }
  knitr::kable(participation_stats, caption = "Participation Statistics")
} else {
  cat("⚠️  Cannot analyze participation patterns - missing clean metrics data\n")
}
```

## Step 8: Visualizing Participation

Create visualizations to understand participation patterns:

```{r visualize-participation}
if (exists("student_summary") && nrow(student_summary) > 0) {
  # Get the name column used in the summary
  name_col <- if ("preferred_name" %in% names(student_summary)) "preferred_name" else "name"
  
  # Plot participation by utterance count
  if ("utterance_count" %in% names(student_summary)) {
    p1 <- ggplot(student_summary, aes(x = reorder(!!sym(name_col), utterance_count), y = utterance_count)) +
      geom_bar(stat = "identity", fill = "steelblue") +
      coord_flip() +
      labs(
        title = "Student Participation by Utterance Count",
        x = "Student Name",
        y = "Number of Utterances"
      ) +
      theme_minimal()
    
    print(p1)
  }
  
  # Plot participation by duration
  if ("total_duration" %in% names(student_summary)) {
    p2 <- ggplot(student_summary, aes(x = reorder(!!sym(name_col), total_duration), y = total_duration / 60)) +
      geom_bar(stat = "identity", fill = "darkgreen") +
      coord_flip() +
      labs(
        title = "Student Participation by Speaking Time",
        x = "Student Name",
        y = "Speaking Time (minutes)"
      ) +
      theme_minimal()
    
    print(p2)
  }
  
  # Save plots
  if (exists("p1")) {
    ggsave("reports/participation_by_utterances.png", p1, width = 10, height = 8)
    cat("Plot saved: reports/participation_by_utterances.png\n")
  }
  if (exists("p2")) {
    ggsave("reports/participation_by_duration.png", p2, width = 10, height = 8)
    cat("Plot saved: reports/participation_by_duration.png\n")
  }
} else {
  cat("⚠️  Cannot create visualizations - missing student summary data\n")
}
```

## Step 9: Identifying Participation Gaps

Look for students who might need encouragement to participate:

```{r identify-gaps}
if (exists("student_summary") && nrow(student_summary) > 0) {
  # Identify students with low participation
  if ("utterance_count" %in% names(student_summary)) {
    median_utterances <- median(student_summary$utterance_count, na.rm = TRUE)
    low_participation <- student_summary %>%
      filter(utterance_count < median_utterances) %>%
      arrange(utterance_count)
  } else if ("total_duration" %in% names(student_summary)) {
    median_duration <- median(student_summary$total_duration, na.rm = TRUE)
    low_participation <- student_summary %>%
      filter(total_duration < median_duration) %>%
      arrange(total_duration)
  } else {
    cat("⚠️  Cannot identify low participation - no suitable metrics available\n")
    low_participation <- student_summary
  }
  
  # Get the name column used in the summary
  name_col <- if ("preferred_name" %in% names(student_summary)) "preferred_name" else "name"
  
  cat("Students with below-median participation:\n")
  if ("total_duration" %in% names(low_participation)) {
    knitr::kable(low_participation[, c(name_col, "utterance_count", "total_duration")], 
                caption = "Students with Below-Median Participation")
  } else {
    knitr::kable(low_participation[, c(name_col, "utterance_count")], 
                caption = "Students with Below-Median Participation")
  }
  
  # Create participation categories
  if ("utterance_count" %in% names(student_summary)) {
    participation_categories <- student_summary %>%
      mutate(
        participation_level = case_when(
          utterance_count >= quantile(utterance_count, 0.75, na.rm = TRUE) ~ "High",
          utterance_count >= quantile(utterance_count, 0.25, na.rm = TRUE) ~ "Medium",
          TRUE ~ "Low"
        )
      )
  } else if ("total_duration" %in% names(student_summary)) {
    participation_categories <- student_summary %>%
      mutate(
        participation_level = case_when(
          total_duration >= quantile(total_duration, 0.75, na.rm = TRUE) ~ "High",
          total_duration >= quantile(total_duration, 0.25, na.rm = TRUE) ~ "Medium",
          TRUE ~ "Low"
        )
      )
  } else {
    participation_categories <- student_summary %>%
      mutate(participation_level = "Unknown")
  }
  
  # Summary by participation level
  level_summary <- participation_categories %>%
    group_by(participation_level) %>%
    summarise(
      count = n(),
      percentage = n() / nrow(participation_categories) * 100
    )
  
  cat("\nParticipation Level Distribution:\n")
  knitr::kable(level_summary, caption = "Participation Level Distribution")
} else {
  cat("⚠️  Cannot identify participation gaps - missing student summary data\n")
}
```

## Step 10: Creating Actionable Insights

Generate insights that can help improve equitable participation:

```{r actionable-insights}
cat("=== RECOMMENDATIONS FOR EQUITABLE PARTICIPATION ===\n")
cat("1. Students with low participation may benefit from:\n")
cat("   - Direct invitations to contribute\n")
cat("   - Smaller group discussions\n")
cat("   - Alternative participation methods (chat, polls)\n\n")

cat("2. Consider implementing:\n")
cat("   - Structured discussion protocols\n")
cat("   - Think-pair-share activities\n")
cat("   - Anonymous participation options\n\n")

cat("3. Monitor participation patterns over time to:\n")
cat("   - Track improvement in engagement\n")
cat("   - Identify effective interventions\n")
cat("   - Ensure all students feel included\n\n")

if (exists("student_summary") && nrow(student_summary) > 0) {
  cat("4. Specific recommendations for your class:\n")
  cat("   - Total students analyzed:", nrow(student_summary), "\n")
  cat("   - Average utterances per student:", round(mean(student_summary$total_utterances, na.rm = TRUE), 1), "\n")
  cat("   - Students with below-average participation:", sum(student_summary$total_utterances < mean(student_summary$total_utterances, na.rm = TRUE)), "\n")
}
```

## Step 11: Final Privacy Validation

Before saving your analysis, let's do a final privacy check to ensure no sensitive data is exposed:

```{r final-privacy-check}
# Function to scan for real names in final outputs
scan_for_real_names <- function(data, description) {
  cat(sprintf("\n=== Final Privacy Scan: %s ===\n", description))
  
  # Convert data to text
  data_text <- paste(capture.output(print(data)), collapse = " ")
  
      # Look for real name patterns
    real_name_pattern <- "\\b[A-Z][a-z]+(\\s+[A-Z][a-z]+)*\\b"
  real_names <- unlist(regmatches(data_text, gregexpr(real_name_pattern, data_text)))
  
  # Filter out common words that might match
  common_words <- c(
    "Test Report", "Real World", "Test Date", "Test Results", "Test Summary",
    "Package Version", "Total Tests", "Success Rate", "Detailed Results",
    "Status Started", "Status Passed", "Status Failed", "Timestamp Details",
    "Error Handling", "Privacy Features", "Recommendations Review"
  )
  real_names <- real_names[!real_names %in% common_words]
  
  if (length(real_names) > 0) {
    cat("🚨 PRIVACY ISSUE: Real names found in final output\n")
    cat("Names found:", paste(unique(real_names), collapse = ", "), "\n")
    cat("This violates FERPA compliance requirements.\n")
    return(FALSE)
  } else {
    cat("✅ Privacy maintained: No real names in final output\n")
    return(TRUE)
  }
}

# Check final outputs for privacy
if (exists("clean_metrics")) {
  scan_for_real_names(clean_metrics, "Clean Metrics")
}

if (exists("student_summary")) {
  scan_for_real_names(student_summary, "Student Summary")
}

cat("\n✅ Final privacy validation completed.\n")
```

## Step 12: Saving Your Analysis

Save your processed data for future reference:

```{r save-analysis}
# Create output directories
if (!dir.exists("outputs")) {
  dir.create("outputs")
}
if (!dir.exists("reports")) {
  dir.create("reports")
}

# Save the clean metrics
if (exists("clean_metrics")) {
  write_metrics(clean_metrics, what = "engagement", path = "outputs/clean_engagement_metrics.csv", comments_format = "text")
  cat("Saved: outputs/clean_engagement_metrics.csv\n")
}

# Save the student summary
if (exists("student_summary")) {
  write_metrics(student_summary, what = "engagement", path = "outputs/student_participation_summary.csv")
  cat("Saved: outputs/student_participation_summary.csv\n")
}

# Save the name mappings
if (exists("custom_mappings")) {
  write_metrics(custom_mappings, what = "engagement", path = "outputs/name_mappings.csv")
  cat("Saved: outputs/name_mappings.csv\n")
}

cat("\nAnalysis files saved to outputs/ directory\n")
cat("Reports and visualizations saved to reports/ directory\n")
```

## Best Practices and Tips

### Privacy and FERPA Compliance
- **Always test privacy levels** before working with real data
- **Use appropriate privacy levels** for your use case:
  - `ferpa_strict` for maximum privacy (recommended for research)
  - `ferpa_standard` for standard educational use
  - `mask` for basic privacy protection
  - `none` only for temporary matching (never for final outputs)
- **Validate privacy compliance** at each step
- **Never commit or share outputs** containing real names
- **Document your privacy approach** for institutional review

### Data Organization
- Keep your transcript files organized by date/session
- Maintain a consistent naming convention
- Back up your original data files
- Document any manual name mappings

### Ethical Considerations
- Use this data to promote equitable participation, not surveillance
- Focus on group patterns rather than individual performance
- Respect student privacy and preferences
- Share insights constructively with students when appropriate

### Common Pitfalls to Avoid
- Don't assume low participation means disengagement
- Consider cultural and personal communication preferences
- Remember that quality of participation matters more than quantity
- Avoid making assumptions about student abilities based on participation

### Interpreting Results
- Look for patterns, not individual judgments
- Consider context (class size, format, topic)
- Use data to inform teaching strategies, not student evaluation
- Focus on creating inclusive learning environments

## Next Steps

This analysis provides a foundation for understanding participation patterns. Consider:

1. **Regular monitoring**: Track participation patterns over the semester
2. **Intervention strategies**: Implement targeted approaches for low-participation students
3. **Student feedback**: Ask students about their participation preferences
4. **Pedagogical adjustments**: Modify teaching strategies based on insights
5. **Continuous improvement**: Refine your analysis approach over time

Remember: The goal is to create an inclusive learning environment where all students feel comfortable and encouraged to participate in ways that work for them. 